### 2015-07-14 manuscript/defs2.tex
Updating defs2.txt

\subsection{REFDs}
In RAIDER, everything is defined in terms of $l$-mers.  hHence we need
to generlize slightly to allow the ambiguity character, so we will
create a description of a {\it repeat element family discriptior},
with an $l$-mer being a special case of that.

@@

\begin{definition}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the
generlized $l$-mer (w.r.t. $s$), or $\gamma_s(r)$, is the refd created by replacing
all basses corresponding to 0s with * characters.
\end{definition}
Example: $\gamma_{110101}(AACGTG) = AA*G*G$.

@@

@@

I want to get a definition of elementary repeats that paralleles the
one we had.  (And, hopefully, is provably equivilent for spaced seeds
without spaces.)  This is tricky: I'm not yet sure what the equivilent
is for the minimality condition (there is no subsequence of an
elementary repeat that occurs independently of the instances of the
elementary repeat).  Lets look why.

Suppose we have in the string out genome $AAAAATCCCCC$ and
$AAAAAGCCCCC$.  Using the seed $s=11011$, we should get the refd
$r=AA*AA*CC*CC$. Here are different ways of phrasing the minimality
condition, and why they break:
\begin{enumerate}
\item No sub-refd can match any string in the genome outside of those
  matching $r$.  Doesn't work: we can have an AAAAT occuring somewhere
  else wihtout breaking this family, since the second $l$-mer is
  irrelevant to construction.
\item There cannot be any string $S'$, $|s'| < |r|$, occuring outside
  the instances, such that $s'$ generalizes to a sub-refd $s$ or $r$.
\end{enumerate}

\begin{definition}
Let $s$ be a fixed seed with length $l$, $f$ be a fixed integer, and $G$ be a fixed
genomic sequence.  An
\underline{Elementary Repeat Family} is a tuple $(r,S)$ where: 
\begin{enumerate}
\item $r$ is an refd such that $s$ covers $r$. 
\item $S$ is a set of genomic coordinates such that $|S| \geq f$.  Say
  that the sub-refds induced by $S$ are the substrings of $G$ starting
  at cordinates of $S$ and extended $|r|$ bases.
\item For every $i \in S$, $r$ is consistent with the string
  $G_{i:i+|r|}$.
\item Minimality condition.  Possibilities (I'm not sure which is
  right):
  \begin{itemize}
  \item There is no sub-refd $r'$ of $r$ that is covered by $s$ and is
     consistent some sequence $x$ that is sits outside the sequences
      induced by $S$.
  \item There is no $l$-mer $r'$ contained in $r$ that is consistence
    with $s$, such that $\sigma_s(r)'$ occurs anywhere in $G$ outside
    of the sequences induced by $S$.
 \end{itemize}
\item There is no refd $r'$ containing $r$ such that every sequence
  induced by $S$ is contained by some sequence consisted with $r'$.
\end{enumerate}
\end{definition}

{\bf Comments:} For the last two points, I'm really not sure if the
substring relation is the correct relation between $r$ and $r'$.

@@

Right now, I'm trying trying to think ofpossibly interesting
statements and prove them.

Assume a fixed seed $s$ of length $l$ and a fixed genome $G$.

>>>>

\subsection{REFDs}
In RAIDER, everything is defined in terms of $l$-mers.  Hence we need
to generlize slightly to allow the ambiguity character, so we will
create a description of a {\it repeat element family discriptior},
with an $l$-mer being a special case of that.

@@

\begin{definition}
\label{generalize}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the
generlized $l$-mer (w.r.t. $s$), or $\gamma_s(r)$, is the refd created by replacing
all basses corresponding to 0s with * characters.
\end{definition}
Example: $\gamma_{110101}(AACGTG) = AA*G*G$.

@@

\begin{definition}
\label{rmdecomp}
Given an refd $r$ and a seed $s$ that covers $R4, the
\underline{right-most} decomposition, denoted $\delta_s(r)$ is the decomposition $S$ created
by:
\begin{enuemrate}
\item Set $i=0$.
\item Let $S = S \cup \{i\}$.
\item If $i == |r| - l$: quit.
\item Let $j = max_{i < j < i+|l|}\{s \mbox{ hits } r \mbox{ at } j$.
\item Goto (1).
\end{enumerate}
(In other words: Add $0$ to $S$, then keep taking then keep adding the
right-most index you can where $s$ hits $r$ and doesn't leave a gap
from the last hit.)

{\bf Note:} I think later we should consder allowing it to leave a
small gap to allow for insertions.

{\bf Examples:}
\begin{itemize}
\item $\delta_{11011}(AA*AATT*TT) = \{0,5\}$.
\item $\delta_{11011}(AA*AA*TT*TT) = \{0,3,6\}$.
\end{itemize}

\begin{observation}
A few observations about the decompositions of $r$ w.r.t. to seed $s$:
\begin{itemize}
\item Any decomposition must containt $0$ and $|r|-|s|$ (as these are
  the only way the first and last base can be covered).
\item $\delta_s(r)$ exists if and only if there is some decompostion of
  $r$ w.r.t $s$, which is true if and only if $s$ covers $r$.
\item $\delta_s(r)$ is unique if it exists.
\end{itemize}

@@

\begin{definition}
Let $s$ be a fixed seed with length $l$, $f$ be a fixed integer, and $G$ be a fixed
genomic sequence.  An
\underline{Elementary Repeat Family} is a tuple $F=(r,s,S)$ where: 
\begin{enumerate}
\item $s$ is a seed and $r$ is an refd such that $s$ covers $r$. 
\item $S$ is a set of genomic coordinates such that $|S| \geq f$.  Say
  that the sequences induced by $S$ are the substrings of $G$ starting
  at cordinates of $S$ and extended $|r|$ bases.
\item For every $i \in S$, $r$ is consistent with the string
  $G_{i:i+|r|}$.
\item Consider the sub-refds of $r$ used for $\delta_r(s)$ (definition
  \ref~{rmdecomp$}). None of these strings may generalize (definition
  \ref~{generalize}) any substring of $G$ outside ofthose induced by
  $S$.  (Minimiality condition.)
\item There is no refd $r'$ such that all of the following are true:
\begin{itemize}
  \item $r$ generalizes to some proper(?) substring of $v$ of $r'$.
  \item The strings in $\delta_r(s)$ make up part of the cover of
    $\delta_{s}(r')$.
  \item Every string induced in $G$ by $S$ is covered by a string that
    generalizes to $r'$.
\end{itemize}
\end{enumerate}
\end{definition}

{\bf Comments:} 
\begin{itemize}
\item I'm not sure if the maximality condition is what we want.
\item I haven't done anything to control the number of * symbols.
  Currently, we can have multipled refds for the same family by
  generalizing.  I'm not sure if this matters.
\end{itemize}

\begin{definition}
Given a repeat family $F=(r,S)$, the length of the family $|F|=|r|$,
and the size of the family $||F||=|S|$, the number of elements in the
family.
\end{definiton}

Alternative definition of elementary repeats:
\begin{definition}
Let $F=(r,s,S)$ be a repeat family with refd $r$, spaced seed $s$, and
coordinate set $S$ such that $s$ covers $r$. Let $C$ be the set of all
sub-refds corresponding to the right-most decompostion of $S$
w.r.t. $r$.  Then $F$ is a repeat element family (w.r.t. to $s$) if:
\begin{enumerate}
\item $|C| > 0$.
\item For each $i \in S$, $G[i:i+|r|]$ generalizes to $r$.
\item For each $w \in C$, the number of strings in $G$ that generlize
  to $w$ is $|S|$.  (Minimality -- but only applies to the
  decomposition.)
\item $F$ is maximal w.r.t to this definition.  Specifically, we
  cannot extend $r$ in either direction and have it covering exactly
  the same sequences.  (Need to think about this.)
\end{enumerate}

@@

Right now, I'm trying trying to think of possibly interesting
statements and prove them.  Mostly just to get into the swing of this.

\begin{defintion} 
Conisider to elementary repeat families $F_1=(r_1,s,S_1)$ and
$F_2=(r_2,s,S_2)$.  These families overlap if there exist
some $i \in F_1$ and $j \in F_2$ such that either $0 \leq j-i < |r_1|$
or $0 \leq i-j < |r_2|$.
\end{definition}

That is, you have one sequence induced by each family such that they
actually overlap on $G$.

\begin{lemma}
Given repeat families $F_1$ and $F_2$ that overlap, and w.l.o.g. 
pick the $i$ and $j$ such that $0 \leq j-i < |r_1|$.  Then there exist
an offset vale $o = j-i$ such that, for any $i' \in S_1$, $j' \ in S_2$
s.t. $j' \geq i'$, either $j' - i' > |r_1|$ or $j' - i' = o$.
\end{lemma}
(In otherwords, corresponding sequences are offset by the same amount
every time.)

\begin{lemma} 
If repeat element families $F_1$ and $F_2$ overlap, then $||F_1|| \neq
||F_2||$.  
\end{lemma}
